---
title: "Homework1"
author: "Kangyu Xu (kx2224)"
date: "2024-09-20"
output: pdf_document
---

## Question1
(a) Qualitative: Ordinal
(b) Qualitative: Binary
(c) Qualitative: Nominal
(d) Quatitative: Continuous
(e) Quatitative: Discrete

## Question2
```{r}
# Basic info
bike_scores <- c(45, 39, 25, 47, 49, 5, 70, 99, 74, 37, 99, 35, 8, 59)
car_scores <- c(67, 50, 85, 43, 64, 35, 47, 97, 58, 58, 10, 56, 50)
```

### a) Compute Mean, Median, Range, SD
```{r}
mean_bike = mean(bike_scores)
median_bike = median(bike_scores)
range_bike = range(bike_scores)
sd_bike = sd(bike_scores)
cat("Mean (bike crash group):", mean_bike, "\n")
cat("Median (bike crash group):", median_bike, "\n")
cat("Range (bike crash group):", range_bike, "\n")
cat("Standard Deviation (bike crash group):", sd_bike, "\n")
```

### b) Description
```{r}
# Box plot
accident_type <- factor(c(rep("Bike", length(bike_scores))))
depression_scores <- c(bike_scores)

boxplot(depression_scores, 
        main = "Depression Scores by Bike Crash",
        ylab = "Depression Scores",
        border = "black")

```
Therefore the boxplot shows that the median of this dataset is 46. And the shape of box shows that the underlying distribution seems like right-skewed and unimodal.

### c) Plot
```{r}
# Box plot
accident_type <- factor(c(rep("Bike", length(bike_scores)), rep("Car", length(car_scores))))
depression_scores <- c(bike_scores, car_scores)

boxplot(depression_scores ~ accident_type, 
        main = "Depression Scores by Type of Accident",
        xlab = "Type of Accident",
        ylab = "Depression Scores",
        col = c("lightblue", "lightgreen"))

```

### d) description
The boxplot of bike-crash dataset shows that the median of this dataset is 46. And the shape of box shows that the underlying distribution seems like right-skewed and unimodal. The boxplot of car-crash dataset shows that the median of this dataset is 58. And the shape of box shows that the underlying distribution seems like symmetric and unimodal.

### e) Comparison
The bike crash group has a lower median depression score (around 46) compared to the car crash group (with a median around 58). This suggests that the typical depression score is lower in bike crash group.

## Question3
### a)
$$ 
\begin{array}{l}
A = \text{{"an even number appears"}} \\ P(A) = \frac{1}{2}
\end{array}
$$

### b)
$$
\begin{array}{l}
B = \text{{"number 10 appears"}}\\
P(B) = \frac{1}{12}
\end{array}
$$

### c)
$$
\begin{array}{l}
\because B \subset A \\
\therefore P(B \cup A) = P(A) = \frac{1}{2}
\end{array}
$$

### d)
A and B are not independent. This is because if A does not happen, B cannot happen neither.

## Question4
$$
\begin{array}{l}
P(dementia) = 5\% \\
P(positive|dementia) = 80\% \\
P(positive|\neg dementia) = 10 \% \\
--------------------------- \\
  \begin{aligned}
  \therefore P(dementia|positive) &= \frac{P(dementia,positive)}{P(positive)}\\
  &= \frac{P(positive|dementia)\times P (dementia)}{\sum_{s \subset\{dementia,\neg dementia\}} P(positive|s)\times P(s)}\\
  &= \frac{0.8*0.05}{0.8*0.05+0.1*0.95}\\
  &= \frac{8}{27}\\
  \end{aligned}
\end{array}
$$
So the probability that she actually has dementia is about 0.2962
